---
title: "HawkEars_conversion_script"
author: "Kevin Kelly"
date: "2024-10-16"
output: html_document
---

# This script will combine all HawkEars outpur files and convert them into a WT tag csv

## 1. Load libraries and enter files location

```{r}
library(stringr)
library(tidyr)
library(dplyr)
library(fs)
#library(tibble)
#library(data.table)
#library(purrr)
#library(readr)
#library(hms)

# Main working directory
root <- "C:/Users/leona/OneDrive/Desktop/UofA/BIOL 399/R Files" # change to match your folder path

#Location of HawkEars outputs within root directory
Hawkears_output <- "HawkEars" # change to match your output folder
```

# Identify and separate 0 kb files

```{r, include=FALSE}
# Step not complete
```

#Read in Hawkears output files within the output folder into a df; append file names in new column; split species and score into separate columns

```{r}

# List only .txt files in the folder
files <- list.files(path = path, pattern = "\\.txt$", full.names = TRUE)

# Initialize a list to store results
files_combined <- list()

# Function to read and clean files by removing null characters
read_clean_file <- function(file) {
  tryCatch({
    # Read the raw bytes of the file
    raw_data <- readBin(file, "raw", file.info(file)$size)
    
    # Convert to character while removing null characters (0x00)
    clean_text <- rawToChar(raw_data[raw_data != as.raw(0x00)])
    
    # Read the cleaned text as a table
    data <- read.table(text = clean_text, header = FALSE, sep = "")
    
    # Return the data
    return(data)
  }, error = function(e) {
    message("Error reading file:", file, " - Skipping.")
    return(NULL)
  })
}

# Loop through files and attempt to read each one
for (i in seq_along(files)) {
  cat("Reading file:", files[i], "\n")  # Print current file for debugging
  data <- read_clean_file(files[i])
  if (!is.null(data)) {
    # Add the file path to the data
    data <- cbind(data, file_path = files[i])
    # Store in the list
    files_combined[[i]] <- data
  }
}

# Filter out NULL entries if any files failed
files_combined <- Filter(Negate(is.null), files_combined)

# Combine all dataframes into one
files_df <- do.call("rbind", files_combined)

# Assign column names
colnames(files_df)[c(1, 2, 3, 4)] <- c("start_time", "end_time", "tag", "file_path")


# Remove files and i objects
rm(files)
rm(i)

# Drop file_path and split tag into species and score
detections_df <- files_df %>%
  mutate(file_name = path_file(file_path), .keep = "unused") %>%
  separate_wider_delim(tag, delim = ";", names = c("species", "score"), cols_remove = FALSE)

```

# Define target species and filter to only include records for the target species

```{r, include=FALSE}
target_spp_list <- c("TEWA", "WTSP", "OVEN") # add all species of interest to list in this object

detections_target <- detections_df %>%
  filter(species %in% target_spp_list)
```

## Optional

# Determine which detections amongst overlapping detections have the higher score and keep those, while dropping those with lower scores 

```{r, include=FALSE}
detections_keep <- detections_target %>% 
  mutate(keep = case_when(
  start_time - lag(start_time) < 3 & file_name == lag(file_name) & as.numeric(score) - lag(as.numeric(score)) < 0 ~ "Drop",
  lead(start_time) - start_time < 3 & file_name == lead(file_name) & as.numeric(score) - lead(as.numeric(score)) <= 0 ~ "Drop",
  TRUE ~ "Keep")) %>%
  filter(keep == "Keep") %>%
  select(!(keep))

```

## Optional

# Write csv of target species detections. Either including all detections, or including only the highest ranked of overlapping detections

```{r}
# All records of target species
write.csv(detections_target, file = str_c(root, "/Output/file_detections_all.csv", sep = ""), row.names = FALSE) # save to chosen folder in root directory

# Only highest ranking overlapping records
write.csv(detections_keep, file = str_c(root, "/Output/file_detections_keep.csv", sep = ""), row.names = FALSE) # save to chosen folder in root directory
```

# Choose and Define thresholds for each species of interest (using AOU codes for species)

```{r}
threshold_df <- data.frame(
  species = c("TEWA", "WTSP", "OVEN"), # change "SPPx" to 4-letter AOU code of species of interest
  threshold = c(0.35, 0.44, 0.55) # change values to the values that correspond to the species, in the same order as the species appear in the df
)
```

# Enter length of recordings to define task length later

```{r}
recordingLength <- 180 # time in seconds of recording length or desired task length in WT

```
# Pare down detections to those above threshold and clean up dataframe

```{r}
detections_threshold <- detections_keep %>%
  left_join(threshold_df, by = "species") %>%
  filter(score >= threshold)%>%
  mutate(recording_id = str_remove_all(file_name, "_HawkEars.txt")) %>%
  mutate(recording_id = as.factor(recording_id)) %>%
  select(!c(file_name, tag, threshold))
```

# Create df for tags export

```{r}
tags_WT <- detections_threshold %>%
  separate(recording_id, into = c("location", "Date", "Time"), sep = "_") %>%
  mutate(recordingDate = as.POSIXct(
      paste0(substr(Date, 1, 4), ":",   # Year
             substr(Date, 5, 6), ":",   # Month
             substr(Date, 7, 8), " ",   # Day
             substr(Time, 1, 2), ":",   # Hour
             substr(Time, 3, 4), ":",   # Minute
             substr(Time, 5, 6)         # Second
             ), format = "%Y:%m:%d %H:%M:%S")) %>%
  mutate(method = "None") %>%
  mutate(taskLength = recordingLength) %>%
  group_by(location, recordingDate, species) %>%
  mutate(speciesIndividualNumber = row_number()) %>% # generates iterative speciesIndividualNumber values 
  ungroup() %>%
  rename(startTime = start_time) %>%
  mutate(tagLength = end_time - startTime, .after = startTime) %>%
  relocate(startTime, .after = speciesIndividualNumber) %>%
  relocate(tagLength, .after = startTime) %>%
  relocate(species, .after = method) %>%
  select(!c(Date, Time, score, end_time))

```

# Export csv file to upload to Project in WT to create tags

```{r}
# Convert recordingDate to a character format with exact date and time
tags_WT$recordingDate <- format(as.POSIXct(tags_WT$recordingDate, format = "%Y-%m-%d %H:%M:%S"), "%Y-%m-%d %H:%M:%S")

# Write to CSV, preserving the exact date-time format
write.csv(tags_WT, str_c(root, "/Output/smoke_project_tags.csv"), row.names = FALSE, quote = TRUE)



```

